---
title: "pset 1 econ 21020"
author: "Matthew Zhao"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description
This RMarkdown file is meant to be a more readable equivalent to the R file by the same name. This corresponds with the first pset of Econometrics. 


Here we open the ggplot2 library and set the seed for reproducibility.
```{r, message=FALSE}
library(ggplot2)
set.seed(1000)
```


## Problem 8
#### Part (a)
Generate a vector of n draws from a standard normal rv
```{r}
n <- 10000
mu <- 0
sigma <- 1
x <- rnorm(n, mu, sigma)
```
Plot a histogram of the draws using ggplot2
```{r}
ggplot() + 
  geom_histogram(aes(x = x), binwidth = 0.16) + 
  ylab("Count") + xlab("x") + 
  theme_classic(base_size = 20) + 
  theme(text = element_text(size = 20,  family="serif"))
```
#### Part (b)
Generate a vector of n draws from a uniform(-1, 1) rv
```{r}
n <- 10000
min_y <- -1
max_y <- 1
y <- runif(n, min_y , max_y)
```
Plot a histogram of the draws using ggplot2
```{r}
ggplot() + 
  geom_histogram(aes(x = y), binwidth = 0.08) + 
  ylab("Count") + xlab("y") + 
  theme_classic(base_size = 20) + 
  theme(text = element_text(size = 20,  family="serif"))
```

## Problem 9
#### Part (a)

#### Part (b)
Define a custom function that returns draws from a Bernoulli rv
```{r}
my_rbernoulli <- function (n, p) {
  x <- ifelse(runif(n,0,1)<=p,0,1)
  return(x)
}
```
Test the custom Bernoulli generator function
```{r}
set.seed(1000)
x <- my_rbernoulli(10000, 0.5)
length(x) == 10000 # should return TRUE
mean(x) # should be a number near 0.5
```
#### Part (c)
Define a custom function that returns draws from a Binomial rv
Test the custom Bernoulli generator function
```{r}
my_rbinomial <- function(n, p, m){
  for (y in 1:n) {
    ifelse(y==1,x <- as.vector(sum(my_rbernoulli(m,p))), 
           x <- append(x,sum(my_rbernoulli(m,p))))
  }
  return(x)
}
```
Test the custom Binomial generator function
```{r}
set.seed(1000)
x <- my_rbinomial(10000, 0.5, 10)
length(x) == 10000 # should return TRUE
mean(x) # should be a number near 5
```





